Probability Distribution Checks

### 2017-03-25 Analysis.Rmd
Diagnostics and Analysis
Update diagnostic plots, check exponential fit rather than lognormal

First draft analysis is complete using `MCMCglmm` with diagnostic plots

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ProjectTemplate)
load.project()
```
# 2015

## Leaf Sheath Blight Severity Model
```{r}

# 2015 model goes here

@@

## Tiller Sheath Blight Severity Model
```{r}
TShB_lmm_2015 <- lmer(TShB_AUDPS ~ WMGT * NRTE + (1 | REP/WMGT),
                      data = AUDPS_2015,
                      REML = FALSE)
summary(TShB_lmm_2015)

plot(TShB_lmm_2015)
qqnorm(resid(TShB_lmm_2015))

# 2016

## Leaf Sheath Blight Severity Model
```{r}
LShB_lmm_2016 <- lmer(LShB_AUDPS ~ WMGT * NRTE + (1 | REP/WMGT),
                      data = AUDPS_2016,
                      REML = FALSE)
summary(LShB_lmm_2016)

plot(LShB_lmm_2016)
qqnorm(resid(LShB_lmm_2016))

## Tiller Sheath Blight Severity Model
```{r}
TShB_lmm_2016 <- lmer(TShB_AUDPS ~ WMGT * NRTE + (1 | REP/WMGT),
                      data = AUDPS_2016,
                      REML = FALSE)
summary(TShB_lmm_2016)

plot(TShB_lmm_2016)
qqnorm(resid(TShB_lmm_2016))

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ProjectTemplate)
load.project()
```

## 2015 Leaf Sheath Blight Severity Model
```{r}

eprior <- list(R = list(V = 1, nu = 0.02),
               G = list(G1 = list(V = 1, nu = 0.02, alpha.V = 1000)))
LShB_lmm_2015 <- MCMCglmm(LShB_AUDPS ~ WMGT * NRTE, ~REP, 
                 data = as.data.frame(AUDPS_2015),
                 verbose = FALSE,
                 prior = eprior,
                 pr = TRUE)

summary(LShB_lmm_2015)

# create data frames for generating diagnostic plots
reps <- data.frame(LShB_lmm_2015$Sol[, c(1, 7:10)])
reps <- melt(reps)

trts <-  data.frame(LShB_lmm_2015$Sol[, 2:6])
names(trts)[names(trts) == "WMGTFLD.NRTE100"] <- "FLD.NRTE100"
names(trts)[names(trts) == "WMGTFLD.NRTE120"] <- "FLD.NRTE120"
trts <- melt(trts)

# Create a dummy x-axis variable for plotting
x <- 1:nrow(LShB_lmm_2015$Sol)

# Diagnostic line plots for replicate
ggplot(data = reps, aes(x = rep(x, 5), y = value, group = variable)) +
  geom_line() +
  theme_tufte() +
  xlab("Iteration Number") +
  ylab(NULL) +
  facet_grid(variable ~ .)

# Posterior distributions for replicate
ggplot(reps[reps$variable != "X.Intercept.", ], aes(x = value, color = variable)) + 
  geom_density() +
  theme_tufte()

# Diagnostic line plots for treatements
ggplot(data = trts, aes(x = rep(x, 5), y = value, group = variable)) +
  geom_line() +
  theme_tufte() +
  xlab("Iteration Number") +
  ylab(NULL) +
  facet_grid(variable ~ .)

# Posterior distributions for treatment
ggplot(trts, aes(x = value, color = variable)) + 
  geom_density() +
  theme_tufte()

# joint distibution of error
rdf <- data.frame(LShB_lmm_2015$VCV)
ggplot(rdf, aes(x = sqrt(REP), y = sqrt(units))) + 
  geom_density2d() + 
  geom_abline(intercept = 0, slope = 1)
```

## 2015 Tiller Sheath Blight Severity Model
```{r}

eprior <- list(R = list(V = 1, nu = 0.02),
               G = list(G1 = list(V = 1, nu = 0.02, alpha.V = 1000)))
TShB_lmm_2015 <- MCMCglmm(TShB_AUDPS ~ WMGT * NRTE, ~REP, 
                 data = as.data.frame(AUDPS_2015),
                 verbose = FALSE,
                 prior = eprior,
                 pr = TRUE)

summary(TShB_lmm_2015)

# create data frames for generating diagnostic plots
reps <- data.frame(TShB_lmm_2015$Sol[, c(1, 7:10)])
reps <- melt(reps)

trts <-  data.frame(TShB_lmm_2015$Sol[, 2:6])
names(trts)[names(trts) == "WMGTFLD.NRTE100"] <- "FLD.NRTE100"
names(trts)[names(trts) == "WMGTFLD.NRTE120"] <- "FLD.NRTE120"
trts <- melt(trts)

# Create a dummy x-axis variable for plotting
x <- 1:nrow(TShB_lmm_2015$Sol)

# Diagnostic line plots for replicate
ggplot(data = reps, aes(x = rep(x, 5), y = value, group = variable)) +
  geom_line() +
  theme_tufte() +
  xlab("Iteration Number") +
  ylab(NULL) +
  facet_grid(variable ~ .)

# Posterior distributions for replicate
ggplot(reps[reps$variable != "X.Intercept.", ], aes(x = value, color = variable)) + 
  geom_density() +
  theme_tufte()

# Diagnostic line plots for treatments
ggplot(data = trts, aes(x = rep(x, 5), y = value, group = variable)) +
  geom_line() +
  theme_tufte() +
  xlab("Iteration Number") +
  ylab(NULL) +
  facet_grid(variable ~ .)

# Posterior distributions for treatment
ggplot(trts, aes(x = value, color = variable)) + 
  geom_density() +
  theme_tufte()

# joint distibution of error
rdf <- data.frame(TShB_lmm_2015$VCV)
ggplot(rdf, aes(x = sqrt(REP), y = sqrt(units))) + 
  geom_density2d() + 
  geom_abline(intercept = 0, slope = 1)
```

## 2016 Leaf Sheath Blight Severity Model
```{r}

eprior <- list(R = list(V = 1, nu = 0.02),
               G = list(G1 = list(V = 1, nu = 0.02, alpha.V = 1000)))
LShB_lmm_2016 <- MCMCglmm(LShB_AUDPS ~ WMGT * NRTE, ~REP, 
                 data = as.data.frame(AUDPS_2016),
                 verbose = FALSE,
                 prior = eprior,
                 pr = TRUE)

summary(LShB_lmm_2016)

# create data frames for generating diagnostic plots
reps <- data.frame(LShB_lmm_2016$Sol[, c(1, 5:8)])
reps <- melt(reps)

trts <-  data.frame(LShB_lmm_2016$Sol[, 2:4])
names(trts)[names(trts) == "WMGTFLD.NRTE180"] <- "FLD.NRTE180"
trts <- melt(trts)

# Create a dummy x-axis variable for plotting
x <- 1:nrow(LShB_lmm_2016$Sol)

# Diagnostic line plots for replicate
ggplot(data = reps, aes(x = rep(x, 5), y = value, group = variable)) +
  geom_line() +
  theme_tufte() +
  xlab("Iteration Number") +
  ylab(NULL) +
  facet_grid(variable ~ .)

# Posterior distributions for replicate
ggplot(reps[reps$variable != "X.Intercept.", ], aes(x = value, color = variable)) + 
  geom_density() +
  theme_tufte()

# Diagnostic line plots for treatments
ggplot(data = trts, aes(x = rep(x, 3), y = value, group = variable)) +
  geom_line() +
  theme_tufte() +
  xlab("Iteration Number") +
  ylab(NULL) +
  facet_grid(variable ~ .)

# Posterior distributions for treatment
ggplot(trts, aes(x = value, color = variable)) + 
  geom_density() +
  theme_tufte()

# joint distibution of error
rdf <- data.frame(TShB_lmm_2015$VCV)
ggplot(rdf, aes(x = sqrt(REP), y = sqrt(units))) + 
  geom_density2d() + 
  geom_abline(intercept = 0, slope = 1)

@@

## Tiller Sheath Blight Severity Model
```{r}

eprior <- list(R = list(V = 1, nu = 0.02),
               G = list(G1 = list(V = 1, nu = 0.02, alpha.V = 1000)))
TShB_lmm_2016 <- MCMCglmm(TShB_AUDPS ~ WMGT * NRTE, ~REP, 
                 data = as.data.frame(AUDPS_2016),
                 verbose = FALSE,
                 prior = eprior,
                 pr = TRUE)

summary(TShB_lmm_2016)

# create data frames for generating diagnostic plots
reps <- data.frame(TShB_lmm_2016$Sol[, c(1, 5:8)])
reps <- melt(reps)

trts <-  data.frame(TShB_lmm_2016$Sol[, 2:4])
names(trts)[names(trts) == "WMGTFLD.NRTE180"] <- "FLD.NRTE180"
trts <- melt(trts)

# Create a dummy x-axis variable for plotting
x <- 1:nrow(TShB_lmm_2016$Sol)

# Diagnostic line plots for replicate
ggplot(data = reps, aes(x = rep(x, 5), y = value, group = variable)) +
  geom_line() +
  theme_tufte() +
  xlab("Iteration Number") +
  ylab(NULL) +
  facet_grid(variable ~ .)

# Posterior distributions for replicate
ggplot(reps[reps$variable != "X.Intercept.", ], aes(x = value, color = variable)) + 
  geom_density() +
  theme_tufte()

# Diagnostic line plots for treatments
ggplot(data = trts, aes(x = rep(x, 3), y = value, group = variable)) +
  geom_line() +
  theme_tufte() +
  xlab("Iteration Number") +
  ylab(NULL) +
  facet_grid(variable ~ .)

# Posterior distributions for treatment
ggplot(trts, aes(x = value, color = variable)) + 
  geom_density() +
  theme_tufte()

# joint distibution of error
rdf <- data.frame(TShB_lmm_2015$VCV)
ggplot(rdf, aes(x = sqrt(REP), y = sqrt(units))) + 
  geom_density2d() + 
  geom_abline(intercept = 0, slope = 1)
```

### 2017-03-25 Probability_distribution_checks.Rmd
Update probability distribution checks to use new data frame names
Diagnostics and Analysis
Update diagnostic plots, check exponential fit rather than lognormal

First draft analysis is complete using `MCMCglmm` with diagnostic plots

Here we check the probability distribution that best fits these data. I'm following a method from http://ase.tufts.edu/gsc/gradresources/guidetomixedmodelsinr/mixed%20model%20guide.html. As you see in that guide, there are several distributions that can be checked, for my purposes normal or lognormal are enough so I won't check the others.

@@

qqp(AUDPS_2015$LShB_rating, "norm")

qqp(AUDPS_2015$LShB_rating, "lnorm")

@@

AUDPS_2015$TShB_rating <- AUDPS_2015$TShB_rating + 1

qqp(AUDPS_2015$TShB_rating, "norm")

qqp(AUDPS_2015$TShB_rating, "lnorm")

@@

qqp(AUDPS_2016$LShB_rating, "norm")

qqp(AUDPS_2016$LShB_rating, "lnorm")

@@

qqp(AUDPS_2016$TShB_rating, "norm")

qqp(AUDPS_2016$TShB_rating, "lnorm")

@@

To read the graphs, we want to select a distribution where most of the points fall between the dotted red lines. In this case, for the 2015 data, the leaf data fit a lognormal distribution best, for all others a normal distribution appears to be fine.

>>>>

Here we check the probability distribution that best fits these data. I'm following a method from http://ase.tufts.edu/gsc/gradresources/guidetomixedmodelsinr/mixed%20model%20guide.html. As you see in that guide, there are several distributions that can be checked, for my purposes normal or exponential are enough so I won't check the others.

@@

f1 <- fitdist(AUDPS_2015$LShB_AUDPS, "norm")
plot(f1)
qqp(AUDPS_2015$LShB_AUDPS, "norm")

f2 <- fitdist(AUDPS_2015$LShB_AUDPS, "exp")
plot(f2)
qqp(AUDPS_2015$LShB_AUDPS, "exp")

@@

f1 <- fitdist(AUDPS_2015$TShB_AUDPS, "norm")
plot(f1)
qqp(AUDPS_2015$TShB_AUDPS, "norm")

f2 <- fitdist(AUDPS_2015$TShB_AUDPS, "exp")
plot(f2)
qqp(AUDPS_2015$TShB_AUDPS, "exp")

@@

f1 <- fitdist(AUDPS_2016$LShB_AUDPS, "norm")
plot(f1)
qqp(AUDPS_2016$LShB_AUDPS, "norm")

f2 <- fitdist(AUDPS_2016$LShB_AUDPS, "exp")
plot(f2)
qqp(AUDPS_2015$LShB_AUDPS, "exp")

@@

f1 <- fitdist(AUDPS_2016$TShB_AUDPS, "norm")
plot(f1)
qqp(AUDPS_2016$TShB_AUDPS, "norm")

f2 <- fitdist(AUDPS_2016$TShB_AUDPS, "exp")
plot(f2)
qqp(AUDPS_2015$LShB_AUDPS, "exp")

@@

To read the graphs, we want to select a distribution where most of the points fall between the dotted red lines. On the y-axis are the observations and on the x-axis the quantiles modelled by the distribution. A perfect distribution fit is represented by the solid red line and the dashed red lines are the confidence intervals of the perfect distribution fit. In this case, for the 2015 data, the leaf data fit a lognormal distribution best or exponential, for all others a normal distribution appears to be fine. Because of the non-normal residuals of the 2015 leaf sheath blight data we'll look at using something that can handle these data like `MCMCglmm`.
